% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Function Call Allocates Multiple Resources}
\label{FunctionCallAllocatesMultipleResources::overview}
``CERT Secure Coding RES30-C'' states
\begin{quote}
Allocating more than one resource in a single statement could result in a memory leak, and this could lead to a denial-of-service attack.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal on each SgFunctionCallExp node. For each node get the expression list of its arguments and check if any such argument expressions are the {\tt new} keyword. If the number of {\tt new} expressions exceeds one then flag an error.

\subsection{Non-Compliant Code Example}
\begin{verbatim}
class A
{
};

class B
{
};

int foo( A *a, B *b )
{
  return 0;
}

int main()
{
  A *a = new A;
  B *b = new B;
  int i = foo( a, b ); //ok...

  return foo( new A, new B ); //bad
}
\end{verbatim}

\subsection{Compliant Solution}
See the call to {\tt foo} above.

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Traverse all SgFunctionCallExp nodes
\item For each node get the list of argument expressions
\item Count the number of {\tt new} keyword argument expressions
\item If the number of {\tt new} keyword argument expressions exceeds one then flag an error.
\item Report all violations.
\end{enumerate}

\subsection{References}
\htmladdnormallink{RES30-C. Never allocate more than one resource in a single statement}{https://www.securecoding.cert.org/confluence/display/cplusplus/RES30-C.+Never+allocate+more+than+one+resource+in+a+single+statement}
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
