% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Sub Expression Evaluation Order}
\label{SubExpressionEvaluationOrder::overview}

This checker detects if there exist, within an expression, sub-expressions that update the same variable.
As the order of evaluation of such expressions is not guaranteed to be left-to-right, any of the sub-expressions can be taken place first.

\subsection{Parameter Requirements}

%Write the Parameter specification here.
None.

\subsection{Implementation}

%Details of the implementation go here.
This checker uses a simple traversal. For every function call statement, the checker examines 1) whether the function call has sub-expressions that update variables and 2) the variables updated are identical.

\subsection{Non-Compliant Code Example}

\begin{verbatim}
int bar(int a, int b);

void foo()
{
  int i = 0;

  i = bar(++i, ++i); // either ++i could be evaluated first
  i = bar((i=3), (i=4)); // no particular order is guaranteed.
}
\end{verbatim}

\subsection{Compliant Solution}

\begin{verbatim}

int bar(int a, int b);

void foo()
{
  int i = 0;

  i = bar(2, 3); // fine
  i = bar((i=2), 3); //fine
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item For each node, check if the node is a function call statement.
\item Examine if the function call has sub-expressions that update variables.
\item If yes, examine further if the variables updated are identical.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
The Programming Research Group, High-Integrity C++ Coding Standard Manual, Item 10.3: ``Do not assume the order of evaluation of operands in an expression."
