% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Uninitialized Definition}
\label{UninitializedDefinition::overview}

% write your introduction
This test ensures that all variables are initialized at their point of
definition. The detector will report all variable declarations without
initializer expressions, except for some special cases:
                                                                                                                                                                                                  
\begin{itemize}
\item Variables declared extern: Such declarations refer to variables defined
elsewhere, initializers are therefore not required at this point.
\item Variables declared static: These variables are automatically initialized
to~0 (of the appropriate type) if no explicit initializer is present.
\item Class member variables: The class constructor is responsible for
initializing such variables.
\item Variables of class type: Class objects are default-initialized if no
explicit initializer expression is present.
\item Variables declared at file (`global') scope: File scope declarations are
implicitly static; they can be changed to extern by an explicit modifier. One
of the above cases will always apply.
\end{itemize}
 

\subsection{Parameter Requirements}

This checker does not require any parameters.

\subsection{Non-Compliant Code Example}

\begin{verbatim}
void f_noncompliant()
{
    int x; // not OK, no initializer
}
\end{verbatim}

\subsection{Compliant Solution}

\begin{verbatim}
struct foo {
    int member; // OK, class member
};

void f_compliant(int n)
{
    int x = n; // OK, initializer present
    static int y; // OK, static
    struct foo st; // OK, class type (has constructor)
    extern int not_here; // OK, extern
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item For each variable declaration without an initializer expression, check
the above criteria.
\item If none of the exceptions apply, generate a diagnostic.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
A reference for this rule is: H.~Sutter, A.~Alexandrescu: ``C++ Coding
Standards'', Item~19: ``Always initialize variables''.
