% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{For Loop Cpp Index Variable Declaration}
\label{ForLoopCppIndexVariableDeclaration::overview}

``ALE3D Coding Standards \& Style Guide'' Item \#6.2 states that
\begin{quote}
C++ loop index variables should be declared in the loop statement. Declaration of a loop index variable in the first clause of the for() statement ensures that its scope is limited to the loop body.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that seeks SgForInitStatements that have NULL declaration statements. These nodes are flagged as violations.

\subsection{Non-Compliant Code Example}
This non-compliant code declares the loop control variable outside the loop statement.

\begin{verbatim}
int main()
{
  int i = 0;
  for( i = 0; i < 100; i++ ){}

  return 0;
} //main()
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution declares the index variable inside the loop statement.

\begin{verbatim}
int main()
{
  for( int i = 0; i < 100; i++ ){}

  return 0;
} //main()
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal visiting all declaration statement nodes.
\item For each declaration statement node, if parent node is for loop initialization statement then flag violation.
\item Report any violations. 
\end{enumerate}

\subsection{References}
Arrighi B., Neely R., Reus J. ``ALE3D Coding Standards \& Style Guide'', 2005.
