% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Single Parameter Constructor Explicit Modifier}
\label{SingleParameterConstructorExplicitModifier::overview}

{\it The Elements of C++ Style} item \#104 states that
\begin{quote}
A compiler can use a single parameter constructor for type conversions. While this is natural in some situations, it might be unexpected in others\ldots you can avoid this behavior by declaring a constructor with the {\tt explicit} keyword.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that finds instances of
SgFunctionDeclaration that are constructors with a single parameter. If these
SgFunctionDeclaration are not modified with the ``explicit'' keyword then
a violation is flagged.

\subsection{Non-Compliant Code Example}
This non-compliant code has a single parameter constructor that is not declared with the {\tt explicit} keyword.

\begin{verbatim}
class Class
{
  int num;
  public:
    Class( int n ){ num = n; }
    int getNum() const { return num; }
}; //class Class
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution declares the single parameter constructor with the {\tt explicit} keyword modifier.

\begin{verbatim}
class Class
{
  int num;
  public:
    explicit Class( int n ){ num = n; }
    int getNum() const { return num; }
}; //class Class
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform a simple AST traversal visiting all function declaration nodes.
\item For each function declaration, if node is constructor then check the size of its parameter list.
\item If the parameter list size of constructor is 1 and is not declared with the {\tt explicit} modifier then flag violation.
\item Report any violations. 
\end{enumerate}

\subsection{References}

Bumgardner G., Gray A., and Misfeldt T. {\it The Elements of C++ Style}. Cambridge University Press 2004.

