% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Variable Name Equals Database Name}
\label{VariableNameEqualsDatabaseName::overview}
For some member function accesses the name of the local variable
that gets assigned the result of the function call should have a
name equal to the first argument. [ALE3D] E.g:
\begin{verbatim}
      real8 *sx = regM->fieldReal("sx") ;
      real8 *syy = regM->fieldReal("sy") ;
\end{verbatim}	  
Where the name of the local variable is not the same as the name in the
database for "syy", but it is the same for "sx". 

This checker will only report the locations of the assign expressions
where this rule is not followed.


\subsection{Parameter Requirements}
The checker takes the name of the class and member function that
on call should assign it's result to a variable with the same name
as the first argument. The arguments are "VariableNameEqualsDatabaseName.ClassName=CLASSNAME"
and "VariableNameEqualsDatabaseName.MemberFunctionName=MEMFUNCNAME".

\subsection{Implementation}
The checker will look for a function call to the member function within
the class that we are looking for. When such a call is found it assumes that
the lhs of the last assign expression is the variable access that we are interested in.
If the name of that variable does not satisfy the rule we report an error.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}
#include <string>
class MixMatmodel{

  public:
    double fieldReal(std::string str){ return 1; };

};

int main(){
    MixMatmodel x;

    int y     = x.fieldReal("test1");
    int z     = x.fieldReal("test1:test2");
    int test2 = x.fieldReal("test1:test2:test3");

};
\end{verbatim}

\subsection{Compliant Solution}

% write your compliant code subsection

\begin{verbatim}
#include <string>
class MixMatmodel{

  public:
    double fieldReal(std::string str){ return 1; };

};

int main(){
    MixMatmodel x;

    int test1 = x.fieldReal("test1");
    int test2 = x.fieldReal("test1:test2");
    int test3 = x.fieldReal("test1:test2:test3");

};
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Travese the AST
\item For each call to the member function we are interested in make sure field name is equal to the string provided as an argument.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
