% \begin{center}
% ********  To be removed later  *************  \newline
% \end{center}

%  Purpose:
% \begin{itemize}
%    \item H. List of references
% \end{itemize}
% \begin{center}
% *********************  \newline
% \end{center}
% \vspace{0.25in}

% Bibliography for thesis (used by \include in thesis.tex

%\chapter{Bibliography}

% \begin{center}
% {\bf (NOTE: The references have not been updated to make them consistant). }
% \end{center}

\begin{thebibliography}{WW}
\small


% 22
\vspace {0.5 in}
\bibitem{ExpressionTemplates1997}
{\bf Bassetti, F., Davis, K., Quinlan, D.:}
"A Comparison of Performance-enhancing Strategies for Parallel
Numerical Object-Oriented Frameworks",
To be published in Proceedings of the first International Scientific Computing in Object-Oriented
Parallel Environments (ISCOPE) Conference, Marina del Rey, California, Dec, 1997 \\
 
% 21
\bibitem{MLB1997}
{\bf Quinlan, D., Berndt, M.:}
"MLB: Multilevel Load Balancing for Structured Grid Applications",
Published in Proceedings of the SIAM Parallel Conference, Minneapolis, MN. March, 1997 \\
 
% 20
\bibitem{OVERTURE1997} 
{\bf Brown, D., Chesshire, G., Henshaw, W., and Quinlan, D.:}
"OVERTURE: An Object-Oriented Software System for Solving Partial
Differential Equations in Serial and Parallel Environments",
Published in Proceedings of the SIAM Parallel Conference, Minneapolis, MN. March, 1997 \\
 
% 19
\bibitem{ParallelAMR1997} 
{\bf Balsara, D., Quinlan, D.:}
"Parallel Object-Oriented Adaptive Mesh Refinement",
Published in Proceedings of the SIAM Parallel Conference, Minneapolis, MN. March, 1997 \\
 
% 18
\bibitem{AMR++DesignPaper1997}
 {\bf Quinlan, D.:}
"AMR++: A Design for Parallel Object-Oriented Adaptive Mesh Refinement",
Published in Proceedings of the IMA Workshop on Structured Adaptive Mesh
Refinement, Minneapolis, MN. March, 1997 \\
 
% 17
%\vspace {0.5 in}
\bibitem{OVERTURE1996}
  {\bf Brislawn, K. D., D. L. Brown, G. S. Chesshire, W. D. Henshaw, K. I. Pao, D.
J. Quinlan, W. J. Rider, and J. S. Saltzman:}
"An Object-Oriented Approach to
Grid Generation and PDE Computations Using Adaptive, Moving, Overlapping
Grids",  Presented at 1996 Parallel Object-Oriented Methods and Applications
Conference, Santa Fe, NM; February 29, 1996, and also at the Fifth International
Conference on Numerical Grid Generation in Computational Field Simulations, Mississippi State,
April 4, 1996.
 
% 16
\bibitem{BradleyBrislawnQuinlanZhangNuri1994}
   {\bf Bradley, Brislawn, Quinlan, Zhang, Nuri:}
  "Wavelet subband coding of computer simulation output using the A++ array class library,"
   Proceedings of Space Earth Science Data Compression Workshop, Snowbird, UT, March 1995.
 
% 15
\bibitem{ParsonsQuinlan1994}
  {\bf Parsons, R. and Quinlan, D.:}
  "A++/P++ Array Classes for Architecture Independent Finite Difference Computations,"
   Proceedings of the Second Annual Object-Oriented Numerics Conference, Sunriver, Oregon, April 1994.
 
\bibitem{ParsonsQuinlan1993}
  {\bf Parsons, R. and Quinlan, D.:}
  "Run-time Recognition of Task Parallelism within the P++ Parallel Array Class Library,"
   Proceedings of the Conference on Parallel Scalable Libraries, Mississippi State, 1993.
 


\bibitem{ANGUS1989}
   {\bf Angus I. G. and Thompkins W. T.:}
   Data Storage, Concurrency, and Portability: An Object Oriented
   Approach to Fluid Dynamics; Fourth Conference on Hypercubes,
   Concurrent Computers, and Applications, 1989.

\bibitem{BSKS9920}
   {\bf Baden, S. B.; Kohn, S. R.: }
   Lattice Parallelism: A Parallel Programming Model for Non-Uniform,
   Structured Scientific Computations; Technical report of University 
   of California, San Diego, Vol. CS92-261, September 1992.

\bibitem{BDLMQD11920}
   {\bf Balsara, D., Lemke, M., Quinlan, D.:} AMR++, a C++ Object Oriented
   Class Library for Parallel Adaptive Mesh Refinement Fluid Dynamics
   Applications, Proceeding of the American Society of Mechanical
   Engineers, Winter Anual Meeting, Anahiem, CA, Symposium on Adaptive,
   Multilevel and Hierarchical Computational Stratagies, 
   November 8-13, 1992.

\bibitem{BSSC2910}
   {\bf Berryman, H.; Saltz, J. ; Scroggs, J.: }
   Execution Time Support for Adaptive Scientific Algorithms on Distributed 
   Memory Machines;
   Concurrency: Practice and Experience, Vol. 3(3), pg. 159-178, June 1991.

\bibitem{CHKE1920}
   {\bf Chandy, K.M.; Kesselman, C.:}
   CC++: A Declarative Concurrent Object-Oriented Programming
   Notation; California Institute of Technology, Report, Pasadena, 1992.

\bibitem{CCRS2910}
   {\bf Chase, C.; Cheeung, A.; Reeves, A.; Smith, M.: }
   Paragon: A Parallel Programming Environment for Scientific Applications
   Using Communication Structures; Proceedings of the 1991 Conference
   on Parallel Processing, IL.
\bibitem{FWFJ3900}
   {\bf Forslund, D.; Wingate, C.; Ford, P.; Junkins, S.; Jackson, J.; Pope, S.:}
   Experiences in Writing a Distributed Particle Simulation Code in C++;
   USENIX C++ Conference Proceedings, San Francisco, CA, 1990.
%\bibitem{HAMC1890}
%   {\bf Hart, L.; McCormick, S.F.:}
%   Asynchronous Multilevel Adaptive Methods for Solving Partial Differential
%   Equations on Multiprocessors: Basic Ideas;
%   Parallel Computing 12, 1989, pg. 131-144.
\bibitem{HPFF}
   {\bf High Performance Fortran Forum: }
   Draft High Performance Fortran Language Specification, Version 0.4,
   Nov. 1992. Available from titan.cs.rice.edu by anonymous ftp.
\bibitem{LJGD0910}
   {\bf Lee, J. K.; Gannon, D.: }
   Object-Oriented Parallel Programming Experiments and Results;
   Proceedings of Supercomputing 91 (Albuquerque, Nov.), IEEE Computer
   Society and ACM SIGARCH, 1991, pg. 273-282.
\bibitem{LEQU1910}
   {\bf Lemke, M.; Quinlan, D.:}
   Fast Adaptive Composite Grid Methods on Distributed Parallel Architectures;
   Proceedings of the Fifth Copper Mountain Conference on Multigrid
   Methods, Copper Mountain, USA-CO, April 1991. Also in Communications in Applied 
   Numerical Methods, Wiley, Vol. 8 No. 9 Sept. 1992.
\bibitem{LEQU1920}
   {\bf Lemke, M.; Quinlan, D.:}
   P++, a C++ Virtual Shared Grids Based Programming
   Environment for Architecture-Independent Development 
   of Structured Grid Applications; Arbeitspapiere der GMD, No. 611, 20 pages,
   Gesellschaft f\"{u}r Mathematik und Datenverarbeitung, 
   St. Augustin, Germany (West), February 1992.
\bibitem{LEQU1921}
   {\bf Lemke, M.; Quinlan, D.:}
   P++, a C++ Virtual Shared Grids Based Programming
   Environment for Architecture-Independent Development 
   of Structured Grid Applications; accepted for CONPAR/VAPP V, September
   1992, Lyon, France; to be published in Lecture Notes in Computer Science,
   Springer Verlag, September 1992.
\bibitem{LMQDKW1993}
   {\bf Lemke, M., Quinlan, D., Witsch, K.:}
   An Object Oriented Approach for Parallel Self Adaptive
   Mesh Refinement on Block Structured Grids, Preceedings of
   the 9th GAMM-Seminar Kiel, Notes on Numerical Fluid Mechanics,
   Vieweg, Germany, 1993.
\bibitem{MCQU1890}
   {\bf McCormick, S., Quinlan, D.:}
   Asynchronous Multilevel Adaptive Methods for Solving Partial Differential
   Equations on Multiprocessors: Performance results;
   Parallel Computing, 12, 1989, pg. 145-156.
\bibitem{MCQU1870}
   {\bf McCormick, S.; Quinlan, D.:} Multilevel Load Balancing,
   Internal Report, Computational Mathematics Group, University of Colorado,
   Denver, 1987.
% \bibitem[Oliver]{oliver}Ian Oliver.\(1993\) {\em Programming Classics:
%    Implementing the World's Best Algorithms.}
%    Englewood Cliffs,N.J.:  Prentice Hall.
\bibitem{PBRW3910}
   {\bf Peery, J.; Budge, K.; Robinson, A.; Whitney, D.:}
   Using C++ as a Scientific Programming Language; Report,
   Sandia National Laboratories, Albuquerque, NM, 1991.
\bibitem{SCHO0910}
   {\bf Schoenberg, R.:}
   M++, an Array Language Extension to C++; Dyad Software Corp., Renton,
   WA, 1991.
\bibitem{STRO0910}
   {\bf Stroustrup, B.:}
   The C++ Programming Language, 2nd Edition; Addison-Wesley, 1991.
\bibitem{AJMY:SC98}
V.~Adve, G.~Jin, J.~Mellor-Crummey, and Q.~Yi.
\newblock High performance fortran compilation techniques for parallelizing
  scientific codes.
\newblock In {\em Proceedings of SC98: High Performance Computing and
  Networking}, Nov 1998.

\bibitem{AM:PLDI98}
V.~Adve and J.~Mellor-Crummey.
\newblock Using integer sets for data-parallel program analysis and
  optimization.
\newblock In {\em Proceedings of the {SIGPLAN} '98 Conference on Programming
  Language Design and Implementation}, Montreal, June 1998.

\bibitem{AK:Book}
R.~Allen and K.~Kennedy.
\newblock {\em Optimizing Compilers for Modern Architectures}.
\newblock Morgan Kaufmann, San Francisco, October 2001.

\bibitem{CFH:CPPSC95}
J.~F. L.~Carter and S.~F. Hummel.
\newblock Efficient multiprocessor parallelism via hierarchical tiling.
\newblock In {\em SIAM Conference on Parallel Processing for Scientific
  Computing}, February 1995.

\bibitem{WL:TOPDS91}
M.~E. Wolf and M.~Lam.
\newblock A loop transformation theory and an algorithm to maximize
  parallelism.
\newblock {\em {IEEE} Transactions on Parallel and Distributed Systems},
  2(4):452--471, Oct. 1991.

\bibitem{Wolfe:Book}
M.~J. Wolfe.
\newblock {\em Optimizing Supercompilers for Supercomputers}.
\newblock The MIT Press, Cambridge, 1989.

\bibitem{MyThesis}
Q.~Yi.
\newblock {\em Transforming Complex Loop Nests For Locality}.
\newblock PhD thesis, Rice University, 2002.

\bibitem{YAK:PLDI00}
Q.~Yi, V.~Adve, and K.~Kennedy.
\newblock Transforming loops to recursion for multi-level memory hierarchies.
\newblock In {\em ACM SIGPLAN Conference on Programming Language Design and
  Implementation}, Vancouver, British Columbia, Canada, June 2000.

\bibitem{YK:LACSI02}
Q.~Yi and K.~Kennedy.
\newblock Improving memory hierarchy performance through combined loop
  interchange and multi-level fusion.
\newblock In {\em LACSI Symposium}, Santa Fe, NM, Oct 2002.

\bibitem{LiaoEffective2009}
Chunhua Liao, Daniel J. Quinlan, Richard Vuduc and Thomas Panas.
\newblock Effective Source-to-Source Outlining to Support Whole Program
Empirical Optimization.
\newblock The 22nd International Workshop on Languages and Compilers for Parallel Computing, Newark, Delaware, USA. October 8-10, 2009.

\bibitem{LiaoExtending2009}
Chunhua Liao, Daniel J. Quinlan, Jeremiah J. Willcock and Thomas Panas,
\newblock Extending Automatic Parallelization to Optimize High-Level
Abstractions for Multicore,
\newblock In Proceedings of the 5th international Workshop on OpenMP:
Evolving OpenMP in An Age of Extreme Parallelism (Dresden, Germany, June 03
- 05, 2009).

\bibitem{LiaoXOMP2010}
Chunhua Liao, Daniel J. Quinlan , Thomas Panas and Bronis de Supinski
\newblock A ROSE-based OpenMP 3.0 Research Compiler Supporting Multiple
Runtime Libraries,
\newblock the 6th International Workshop on OpenMP (IWOMP), June 14-16,
2010, Tsukuba, Japan.




\bibitem{YK:Report}
Q.~Yi and K.~Kennedy.
\newblock Transforming complex loop nests for locality.
\newblock Technical Report TR02-386, Computer Science Dept., Rice University,
  Feb. 2002.

\bibitem{UPC:Web}  
\newblock Unified Parallel C.
\newblock http://upc.gwu.edu/

\bibitem{OpenMP3.0}
\newblock OpenMP Application Program Interface Version 3.0
\newblock http://www.openmp.org/mp-documents/spec30.pdf

\bibitem{BUPC:Web}  
\newblock Berkely UPC - Unified Parallel C.
\newblock http://upc.lbl.gov/publications/

\bibitem{BUPC3.10}  
\newblock The Berkeley UPC Runtime Specification V 3.10
\newblock http://upc.lbl.gov/docs/system/upcr.txt

\bibitem{ItaniumCppABI}
\newblock Itanium C++ ABI Revision: 1.86
\newblock http://www.codesourcery.com/public/cxx-abi/abi.html

\bibitem{FADA}
Denis Barthou
\newblock Array Dataflow Analysis in Presence of Non-Affine Constraints.
\newblock PhD thesis, University of Versailles Saint-Quentin en Yvelins, 1998

\bibitem{FADAref}
M.~Belaoucha and D.~Barthou and S.~Touati
\newblock Dataflow Analysis for Irregular Programs. FADA Toolkit User's Guide. Oct, 2009

\bibitem{FADAweb}
Fuzzy Array Dataflow Analysis library.
\newblock http://www.prism.uvsq.fr/~bem/fadalib/

\bibitem{Feautrier}
Paul Feautrier
\newblock Dataflow analysis of array and scalar references.
\newblock Parrallel Processing Letters, 1991.

\bibitem{Rajopadhye}
Sanjay Rajopadhye
\newblock Dependence Analysis and Parallelizing Transformation

\bibitem{PPL}
Parma Polyhedral Library
\newblock http://www.cs.unipr.it/ppl/

\bibitem{PIP}
PipLib
\newblock http://www.piplib.org/

\end{thebibliography}





