\chapter{ROSE Exam Introduction}
\label{rose_exam_introduction}

   There are many subtle aspects of using ROSE, and yet there is a general
attempt to make ROSE as easy to use as possible.  This exam covers
a range of subjects within the use of ROSE and is designed to help
users understand what knowledge base is required to use ROSE well.

It is not critical to take the exam as much as learn from the exam 
by making sure that you understand the questions and answers.
\fixme{Should we attach the answers at the end of each question?}

The purpose of this exam is to test or cover some of the most important parts
of using ROSE and make sure that a number of fundamental parts of ROSE are
understood.  In most cases there are many ways to acomplish the same
result using ROSE and so our goal is to in each case present the most
eligant approaches so that we can provide a foundation for more advanced
use of ROSE (i.e. on much harder problems and in combination with different 
techniques).

Each exam covers analysis and transformations and in some cases on
both source code and binary executables.  Either section (source
code or binary handling) may be skipped as part of focusing on 
ROSE features that are most important.  Each exam covers
a similar topic from both the source code and binary perspective.
Note that since an AST can process both the source code and binary
(both in the same AST) the source code and binary exams can be combined 
into the same exercise, however this way of presenting ROSE is unnecessarily
confusing so the two subjects are handled separately within the exam.

