% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Size Of Pointer}
\label{SizeOfPointer::overview}

% write your introduction
Do not take the size of a pointer to a type when you are trying to
determine the size of the type. Taking the size of a pointer to a type
always returns the size of the pointer and not the size of the type.

This can be particularly problematic when tyring to determine the size
of an array.

\subsection{Parameter Requirements}

%Write the Parameter specification here.
   No Parameter specifications yet!

\subsection{Implementation}

%Details of the implementation go here.

   Finds calls to the sizeof function.  Checks argument for level of
pointer (ie pointer to pointer etc...) then checks the number of
dereference levels.  If these do not match an alert is raised.


\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection
This non-compliant code example mistakenly calls the sizeof() operator
on the variable d\_array which is declared as a pointer to double
instead of the variable d which is declared as a double.
\begin{verbatim}
double *d_array;
size_t num_elems;
/* ... */

if (num_elems > SIZE_MAX/sizeof(d_array)){
  /* handle error condition */
}
else {
  d_array = malloc(sizeof(d_array) * num_elems);
}
\end{verbatim}
The test of num\_elems is to ensure that the multiplication of
sizeof(d\_array) * num\_elems does not result in an integer overflow

For many implementaion, the size of a pointer and the size of double
(or other type) is likely to be different. On IA-32 implementations,
for example, the sizeof(d\_array) is four, while the sizeof(d) is
eight. In this case, insufficient space is allocated to contain an
array of 100 values of type double.

\subsection{Compliant Solution}

% write your compliant code subsection
Make sure you correctly calculate the size of the element to be
contained in the aggregate data structure. The expression sizeof
(*d\_array) returns the size of the data structure referenced by
d\_array and not the size of the pointer.
\begin{verbatim}

double *d_array;
size_t num_elems;
/* ... */

if (num_elems > SIZE_MAX/sizeof(*d_array)){
  /* handle error condition */
}
else {
  d_array = malloc(sizeof(*d_array) * num_elems);


\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Write your checker algorithm
\end{enumerate}

\subsection{References}

% Write some references
\htmladdnormallink{ISO/IEC
  9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References}
[Viega 05] Section 5.6.8, ``Use of sizeof() on a pointer type''
[ISO/IEC 9899-1999] Section 6.5.3.4, ``The sizeof operator''
[Drepper 06] Section 2.1.1, ``Respecting Memory Bounds''
