% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{For Loop Construction Control Stmt}
\label{ForLoopConstructionControlStmt::overview}

``ALE3D Coding Standards \& Style Guide'' item \#6.1 states that
\begin{quote}
for() construction loops must only include statements that control the loop. In particular, for() loops must not initialize or increment/decrement variables not directly related to the loop control.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs the source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that seeks out for loop statement constructs. A list of variables set in the initialization block is generated.  A list of variables set in the increment/decrement block is generated.  Any variable in the increment/decrement list of variables must be in the initialization list of variables.

\subsection{Non-Compliant Code Example}
This non-compliant code initializes the array inside the for() control statement.

\begin{verbatim}
#include <stdlib.h>

int main()
{
  int *array = (int*)malloc( 100*sizeof(int) );
  int j=100;

  for( int i = 0; i < 100; i++, j-- ){
     array[i] = j;
  }

  free(array);
  return 0;
} //main()
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution simply moves the array initialization inside the for() loop body.

\begin{verbatim}
#include <stdlib.h>

int main()
{
  int *array = (int*)malloc( 100*sizeof(int) );

  for( int i = 0; i < 100; i++ ){
     array[i] = i;
  }

  free(array);
  return 0;
} //main()
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal visiting all for statements.
\item Generate a list of variables set in the initialization block.
\item Generate a list of variables set in the increment/decrement block.
\item Report any variable in the increment/decrement list of variables that are not in the initialization list of variables as a violation
\end{enumerate}

\subsection{References}

Arrighi B., Neely R., Reus J. ``ALE3D Coding Standards \& Style Guide'', 2005.
