% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Enum Declaration Namespace Class Scope}
\label{EnumDeclarationNamespaceClassScope::overview}

{\it The Elements of C++ Style} item \#79 states that
\begin{quote}
To avoid symbolic name conflicts between enumerators and other global names, nest enum declarations within the most closely related class or common namespace.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that locates nodes that are enumeration declarations. If a enumeration declaration is found then its parent nodes are traversed until a class or namespace declaration is found. If no namespace or class declaration(s) are found then a violation is flagged by this checker.

\subsection{Non-Compliant Code Example}
This non-compliant code contains an enum declaration at the global scope.

\begin{verbatim}
enum violation{ E1=0, E2, E3 }; // This is a violation
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution simply nests the violation enum declaration in a unique namespace.

\begin{verbatim}
namespace Namespace
{
  enum violation{ E1=0, E2, E3 }; // This is OK
} //namespace Namespace
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform an AST traversal visiting enum declaration nodes.
\item For each enum declaration node visit its parents checking them to be either namespace declarations or class declarations. If no class or namespace declaration parent node is found, then flag violation.
\item Report any violations.
\end{enumerate}

\subsection{References}

Bumgardner G., Gray A., and Misfeldt T. {\it The Elements of C++ Style}. Cambridge University Press 2004.
