% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Const Cast}
\label{ConstCast::overview}

Casting the constness away should never be done.

Casting away constness via const\_cast is just plain false advertising.
If a member function's signature is void someFunc(const foo\& arg); then
the function advertises to it's clients that it will not call any 
non-const member functions on the arg. Casting the constness of arg away
to allow the use of non-const member functions can create unexpected results
for clients of this function.

% write your introduction

\subsection{Parameter Requirements}

%Write the Parameter specification here.
No paramaters are needed.

\subsection{Implementation}
   The checker inspects every cast in the AST. If the type casted to is equal to the type
   casted from minus the const-modifier it is an error.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}
   void foo(){
      const int x = 2;
      int y = (int) x;
   }
\end{verbatim}

\subsection{Compliant Solution}

% write your compliant code subsection

\begin{verbatim}
   void foo(){
      int x = 2;
      int y = x;
   }

\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Traversee the AST
\item If a cast expression casts away constness report an error.
\end{enumerate}

\subsection{References}
ALE3D Style Guide section 14.4.
% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
