\documentclass[12,twoside,openright]{report}
\usepackage{html}

\title{Parallel Ocean Program (POP) User Guide}

\author{Version 2.0 \\
        Los Alamos National Laboratory \\
        LACC 99-18}

\date{23 March 2003}
\begin{document}
\maketitle

\begin{abstract}

This manual contains detailed instructions for operating the
POP code. It should be the only manual needed to use POP ``as is''.
Topics include:

\begin{itemize}
\item How to compile POP, including compile-time options

\item How to run POP, including run-time options in {\em namelist}
      input

\item Procedures for preparing auxiliary input files that are
      needed if you are setting up a new grid

\item Options for model diagnostics

\item Options for model output files and formats
\end{itemize}

Throughout this manual, it is assumed that the operating
system is some variant of Unix.  However, POP has been run
on
\hyperref{PCs under windows.}
         {PCs under windows (see Sec. }{ for details).}
         {sec:PCPOP}.

To keep the {\em User's Guide} as short as possible, a
{\em Reference Manual} is provided for those users who
seek an in-depth understanding of the code, such as
would be needed to make changes to the source code. It
contains a thorough description of the code, including
the model equations, their discretized form, and the
numerical methods used for their solution.  The principal
modules and subroutines are explained, the organization
of the code is outlined, and the principal variables are
defined.

Whenever POP is run, the version number and release date of
the source code on your local system is output at the beginning
of the {\em log file}.  They
should agree with the information on the front cover of this
manual.  If they do not, you can obtain the current version
of this manual at

\noindent
\htmladdnormallink{http://climate.acl.lanl.gov/models/pop/documentation}
                  {http://climate.acl.lanl.gov/models/pop/documentation}.

The Manual revision letter and release date indicate corrections
or additions to the manual corresponding to the stated version
of the model.

If you obtained POP directly from LANL, technical support is
provided by
John Davis (
\htmladdnormallink{jfd@lanl.gov}
                  {mailto:jfd@lanl.gov}, 505-667-4793)
and Phil Jones (\htmladdnormallink{pwjones@lanl.gov}
                {mailto:pwjones@lanl.gov}
                , 505-667-6386).
If you are using POP as a component of the NCAR Climate System Model
(CCSM), technical support is provided by NCAR.

We are always interested in learning about your experiences
with POP, favorable or otherwise. Send your comments and
suggestions for improvements to
\htmladdnormallink{pop\_feedback@acl.lanl.gov}
                  {mailto:pop_feedback@acl.lanl.gov}.
Bug reports should be sent to
\htmladdnormallink{pop\_bugs@acl.lanl.gov}
                  {mailto:pop_bugs@acl.lanl.gov}.

\end{abstract}

\tableofcontents
\listoffigures
\listoftables

\input{POPusers_chap1}
\input{POPusers_chap2}
\input{POPusers_chap3}
\input{POPusers_chap4}
\input{POPusers_chap5}
\input{POPusers_chap6}
\input{POPusers_chap7}

\end{document}
