% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Assignment Operator Check Self}
\label{AssignmentOperatorCheckSelf::overview}
% write your introduction
This test checks to make sure that the first statement in a assignment
operator is a check for self-assignment.  
As noted in An Abbreviated C++ Code Inspection Checklist 12.1.3 .
This will save time allocating new memory and (hopefully) deleting the
 previous copy.  The check for return this; is handled by another checker. 

\subsection{Parameter Requirements}

%Write the Parameter specification here.
No parameters required.

\subsection{Implementation}
  This test checks to make sure that the first statement in a
  assignment operator is a check for self-assignment.  As noted in
 An Abbreviated C++ Code Inspection Checklist 12.1.3 .  This will
 save time allocating new memory and (hopefully) deleting the previous
 copy.  The check for return this; is handled by another checker.
  
%Details of the implementation go here.


\subsection{Non-Compliant Code Example}

\begin{verbatim}

class bike
{
public:
 const bike& operator= (const bike& other);
};

const bike& bike::operator= (const bike& other)
{
...
return *this;
}

\end{verbatim}

\subsection{Compliant Solution}

\begin{verbatim}

const bike& bike::operator= (const bike& other)
{
if (this == &other)
    {return *this;
    }
...
return *this;
}

\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Identify member function
\item Check name for operator=
\item Check first statement as If Statement
\item Check arguments to expression to be this and the right hand argument.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC
% 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
\htmladdnormallink{Abbreviated Code Inspection Checklist}{http://www.lore.ua.ac.be/Teaching/SE3BAC} Section 12.1.3, Assignment Operator''
