% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Fopen Format Parameter}
\label{FopenFormatParameter::overview}
``CERT Secure Coding FIO11-A'' states
\begin{quote}
The C standard specifies specific strings to use for the {\tt mode} for the function {\tt fopen()}. An implementation may define extra strings that define additional modes, but only the modes in the following table (adapted from the C99 standard) are fully portable and C99 compliant:
\end{quote}

\begin{enumerate}
\item r
\item w
\item a
\item rb
\item wb
\item ab
\item r+
\item w+
\item a+
\item r+b or rb+
\item w+b or wb+
\item a+b or ab+
\end{enumerate}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that visits all function call expressions. For each function call expression the function name is confirmed to be {\tt fopen()} then the format parameter is checked against the list of specified strings. If the given parameter is not a standard format string then a violation is flagged.

\subsection{Non-Compliant Code Example}

\begin{verbatim}
#include <stdio.h>

int main()
{
  FILE *f = fopen( "/tmp/tmp.txt", "wr" );

  fclose( f );

  return 0;
}
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution uses the ``r+'' specified parameter string instead.

\begin{verbatim}
#include <stdio.h>

int main()
{
  FILE *f = fopen( "/tmp/tmp.txt", "r+" );

  fclose( f );

  return 0;
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal visiting all function call expression nodes.
\item For each function call expression node, unparse node string then to determine the function name and parse the format parameter.
\item Check the format parameter against list of standard values. If given format parameter does not conform to list of specified values, then flag violation.
\item Report any violations.
\end{enumerate}

\subsection{References}

\htmladdnormallink{Secure Coding : FIO11-A. Take care when specifying the mode parameter of fopen()}{https://www.securecoding.cert.org/confluence/display/seccode/FIO11-A.+Take+care+when+specifying+the+mode+parameter+of+fopen\%28\%29}
