% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{CERT-DCL31-C: Function Definition Prototype}
\label{FunctionDefinitionPrototype::overview}
``CERT Secure Coding DCL31-C'' states
\begin{quote}
Functions should always be declared with the appropriate function prototype. If a function prototype is not available, the compiler cannot perform checks on the number and type of arguments being passed to functions. Argument type checking in C is only performed during compilation, and does not occur during linking, or dynamic loading.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal of function declaration nodes. For each function declaration node find the first non-defining function declaration; if none is found, then flag violation.

\subsection{Non-Compliant Code Example}
This example {\tt foo()} has no prototype.

\begin{verbatim}
int foo( int i )
{
  return i;
}

int main()
{
  return foo(0);
}
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution simply adds a function prototype for {\tt foo()}

\begin{verbatim}
int foo(int);

int foo( int i )
{
  return i;
}

int main()
{
  return foo(0);
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform a simple AST traversal of code visiting all function declaration nodes.
\item For each function declaration node, find first non-defining declaration. If no non-defining declaration is found, then flag violation.
\item Report any violations detected.
\end{enumerate}

\subsection{References}

\htmladdnormallink{Secure Coding : DCL31-C. Ensure every function has a function prototype}{https://www.securecoding.cert.org/confluence/display/seccode/DCL31-C.+Ensure+every+function+has+a+function+prototype}
