% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Cpp Calls Setjmp Longjmp}
\label{CppCallsSetjmpLongjmp::overview}

{\it The Elements of C++ Style} state that:
\begin{quote}
[The setjmp() and longjmp()] functions provide exception handling for C programs. You cannot safely use these functions in C++ code because the exception-handling mechanism they implement does not respect normal object lifecycle semantics--a jump will not result in destruction of scoped, automatically allocated objects.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that seeks out calls
to setjmp() and longjmp() in source files without the ``.c'' extension. These
nodes are flagged as violations.

\subsection{Non-Compliant Code Example}
This contrived trivial example calls setjmp() and longjmp() in C++ code

\begin{verbatim}
#include <setjmp.h>

int main()
{
  jmp_buf env;
  my_container c1;
  int i = setjmp(env);

  if( i != 0 ) exit( 1 );

  int err = c1.clear();

  if( err != 0 ) longjmp( env, 1 );

  return 0;
}
\end{verbatim}

\subsection{Compliant Solution}

The compliant solution uses C++ exception handling.

\begin{verbatim}
int main()
{
  my_container c1;

  try
  {
    c1.clear();    
  }
  catch(...)
  {
    exit( 1 );
  } 

  return 0;
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal visiting function reference codes.
\item Flag all function references named {\tt setjmp()} or {\tt longjmp()} as violations.
\item Report all violations.
\end{enumerate}

\subsection{References}

Bumgardner G., Gray A., and Misfeldt T. {\it The Elements of C++ Style}. Cambridge University Press 2004.
