% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{No Second Term Side Effects}
\label{NoSecondTermSideEffects::overview}
The logical AND and logical OR operators (\&\&, $||$) exhibit ``short
circuit'' operation. That is, the second operand is not evaluated if
the result can be deduced solely by evaluating the first
operand. Consequently, the second operand should not contain side
effects because, if it does, it is not apparent if the side effect occurs

\subsection{Parameter Requirements}

%Write the Parameter specification here.
 No parameters required.

\subsection{Implementation}

%Details of the implementation go here.
We check for And or Or.  We then query for any of a set of operators
known to have side effects.  This checker has the known deficiency of
not checking function calls for side-effects.  To avoid false
positives, it does not notify of functions at all.  

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}
int i;
int max;

if ( (i >= 0 && (i++) <= max) ) {
  /* code */
}
\end{verbatim}

It is unclear whether the value of i will be incremented as a result
of evaluating the condition.

\subsection{Compliant Solution}
In this compliant solution, the behavior is much clearer.

\begin{verbatim}
int i;
int max;

if ( (i >= 0 && (i + 1) <= max) ) {
  i++;
  /* code */
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Find the And or Or operator
\item query the right-hand child for known side-effect having operators
\end{enumerate}

\subsection{References}

% Write some references
\htmladdnormallink{ISO/IEC
  9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References}
[ISO/IEC 9899-1999] Section 6.5.13, ``Logical AND operator,'' and
Section 6.5.14, ``Logical OR operator''
