% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Constructor Destructor Calls Virtual Function}
\label{ConstructorDestructorCallsVirtualFunction::overview}

{\it C++ Coding Standards}, states that:
\begin{quote}
Virtual functions only ``virtually'' always behave virtually: Inside constructors and destructrors, they don't. Worse, any direct or indirect call to an unimplemented {\it pure virtual} function from a constructor or destructor results in undefined behavior. If your design wants virtual dispatch into a derived class from a base class constructor or destructor, you need other techniques such as post-constructors.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file

\subsection{Implementation}
This pattern is checked using a nested AST traversal in which the top level
traversal seeks out definitions of constructors and destructors and two
nested traversals seek out calls to virtual functions of member functions
and non-member functions respectively.

\subsection{Non-Compliant Code Example}
The following code calls a virtual function from a public constructor. This is
a contrived trivial example.

\begin{verbatim}
class Class
{
  int n;

  public:
    Class() { n = Classy(); } //constructor
    ~Class() {} //Destructor

    virtual int Classy() { return 1; }
}; //class Class

int main()
{
  Class c;
  return 0;
} //main()
\end{verbatim}

\subsection{Compliant Solution}

\begin{verbatim}
class Class
{
  int n;

  public:
    Class() { n = 1; } //constructor
    ~Class() {} //Destructor
}; //class Class

int main()
{
  Class c;
  return 0;
} //main()
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform a AST traversal visiting class constructors and destructors.
\item Flag any calls to virtual functions in class constructor or destructor nodes.
\item Report any violations. 
\end{enumerate}

\subsection{References}

Alexandrescu A. and Sutter H. {\it C++ Coding Standards 101 Rules, Guidelines, and Best Practices}. Addison-Wesley 2005.
