%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Using Compass Verifier}

	Compass Verifier is a tool extended from Compass to analyze the source
code of Compass and its checkers to detect certain properties. The motivation
and design of this tool are discussed in Chapter 2 (Design and Verification).
This chapter will detail the Make rules written to run Compass 
Verifier and to setup the parameters to the AllowedFunctions checker.

The rules to setup and run Compass Verifier are found in 
{\tt projects/compass/tools/compassVerifier/Makefile.am}. Follows is an 
overview of their usage labels, ({\tt make oneBigVerify, verify, ...}).
Verifier is designed to examine the compass tool referenced in the environment
variable {\tt \$(TOOLBUILD)} which by default is 
{\tt projects/compass/tools/compass}. This variable may be changed in the
{\tt Makefile.am} of the source directory or on the command line to {\tt make}
in the build tree.
%
\begin{itemize}
\item {\tt oneBigVerify:} a fast but rough static analysis over the Compass
	checker sources. This rule executes {\tt compassVerifier} over the
	concatenation of all sources with names matching those checkers listed
	in {\tt \$(TOOLBUILD)/CHECKER\_LIST\_WITHOUT\_COMMENTS} with extension 
	({\tt .C}).
	These should be the only sources from the checker subdirectories that
	are built with Compass. The consistency of this concatenation depends
	on no global (function, variable, etc.) declaration conflicts; usually,
	this is perserved by default as all checkers employ their own namespace.
%
\item {\tt verify:} a slow and more complete analysis over all files found in
	the checker directories listed in 
	{\tt \$(TOOLBUILD)/CHECKER\_LIST\_WITHOUT\_COMMENTS}.
	The individual rule labels for verify are generated in 
	{\tt verify.makefile} identically to those lower case names found in
	{\tt \$(TOOLBUILD)/CHECKER\_LIST\_WITHOUT\_COMMENTS}. 
	Each rule logs the standard output and error to files suffixed 
	{\tt \_out.txt} and {\tt \_err.txt}
	respectively and the measure of a passing checker is an empty 
	standard error log file. Concurrency may be used with
	({\tt -j\#}) option to GNU Make to reduce the time to complete verify.
%
\item {\tt new\_allow\_list:} regenerates the parameter file 
	{\tt projects/compass/extensions/checkers/allowedFunctions/compass\_parameters} in the
	source tree used as the allowed functions list in Compass Verifier.
	The present list assumes that all sources found in the present source
	tree of Compass are trusted in their use of function calls and whatever
	functions are referenced are allowed. Sources processed by Compass
	Verifier to build the list of allowed functions consist of
	compass.C, compassTestMain.C, buildCheckers.C, and the concatenation
	of checker sources identical to that of {\tt oneBigVerify}. Please
	see the documentation for allowedFunctions
	~\ref{AllowedFunctions::overview} for the details of how
	the new list of allowed functions is created and maintained.
	{\em Not implemented: we plan to limit the recursion of function
	references to exclude those functions called by calls to library
	functions. This should greatly reduce the number of allowed functions
	such that an individual may inspect the generated file.}
\end{itemize}

These three Make rules describe the basic user interface for Compass Verifier.
Other parameters such as those for ForbiddenFunctions should be changed
manually. And an individual should examine the results of Compass Verifier
to ensure a checker meets the validation standards sought-after.
