% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Type Typedef}
\label{TypeTypedef::overview}

Typedefs can be used instead of a type in the cases where it would
increase maintainability or ensure cross-platform compatilibily.

For instance the ALE3D style guide section 3.5 says that the typedef
'Real8' should be used instead of 'double'. The introduction of Real8
is due to the days when a 'double' on CRAY Y-MP meant 16-bytes for
that particular compiler. Although virtually all compilers now interprets
'double' as 8-byte ALE3D use the typedef for consistency.

% write your introduction

\subsection{Parameter Requirements}
Through the 'TypeTypedef.RulesFile' paramaters a filename is provided,
like e.g 'TypeTypedef.RulesFile=/path/to/file/exampleRules.sn'.

The rules are provided through a file where each word is a rule except
everything on the line after a '\#' which is considered a comment. Comments 
can be placed anywhere. There are two types of rules; 'typedefName1',
'c:typedefName2'. The type of the typedef is found in the AST so the user
does not have to type it in. Only declarations with the exact type of the
typedef which does not used the typedef is flagged as an error. When a 
typedef rule is prefixed with 'c:' exact type
matches with and without const modifiers are both allowed.

\subsection{Implementation}

The implementation checks the type of all typed constructs that has the
same type as the typedef and if the typedef is not used it reports an error.
\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}
typedef double Real8;
void foo(){
   double bar; //A violation
};
\end{verbatim}

\subsection{Compliant Solution}

% write your compliant code subsection

\begin{verbatim}
typedef double Real8;
void foo(){
   Real8 bar; //A violation
};
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Traverse the AST
\item For each typed contruct of the same type as the typedef it reports an error if the typedef is not used.
\end{enumerate}

\subsection{References}

ALE3D style guide section 3.5.
% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
