% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Possibly Replicated Variables}
\label{PossiblyReplicatedVariables::overview}

When converting an MPI program to run on a shared-memory (or multi-core)
system, it is important to reduce the amount of memory used by each
process.  One way to do this is to remove data that is independent of the
particular process, and that is only read by the program.  Equation of
state tables in physical simulations are a common example of this: data is
loaded at the beginning of the program, and not modified thereafter.  The
Possibly Replicated Variables analysis suggests variables in the program
that could contain replicated data and thus are good candidates to be
shared between threads on the same system.  Sharing of data using MPI
operations is not considered by this analysis; the only shared data found
is that completely independent of the MPI rank number.

\subsection{Parameter Requirements}

This analyzer does not accept any parameters.

\subsection{Implementation}

The shared variable analysis is implemented using the system dependence
graph (SDG) in ROSE.  It operates on the assumption that the only
information that varies between ranks is the result of calling {\tt
MPI\_Comm\_rank()}; other functions can be added as needed.  It first locates
SDG nodes corresponding to the second argument (pointer to rank) of {\tt
MPI\_Comm\_rank()}, and then traces the uses of those values.  All variables
mentioned in statements using the rank value are considered to be private
to each rank (this is a gross overapproximation of what must be private);
and all values derived from those statements are also considered to be
different across ranks.  All variables of either pointer or array type
(other than parameters and structure fields) in the program are considered
to be eligible for sharing across processes, unless some reason is found to
make them private.  Both the sharable and private variables are reported as
outputs of the analysis.

\subsection{Code Example}

The following code contains both replicated and non-replicated variables:

\begin{verbatim}
int main(int argc, char** argv) {
  int rank, a1[1], a2[1];
  MPI_Init(&argc, &argv);
  MPI_Comm_rank(MPI_COMM_WORLD, &rank);
  a1[0] = 3;    // This variable can be shared among ranks
  a2[0] = rank; // This variable cannot be shared
  printf("a1=%d a2=%d\n", a1[0], a2[0]);
  MPI_Finalize();
  return 0;
}
\end{verbatim}

% \subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
