% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{No Variadic Functions}
\label{NoVariadicFunctions::overview}

``CERT Secure Coding DCL33-C.'' states
\begin{quote}
A variadic function – a function declared with a parameter list ending with ellipsis (...) – can accept a varying number of arguments of differing types. Variadic functions are flexible, but they are also hazardous. The compiler can't verify that a given call to a variadic function passes an appropriate number of arguments or that those arguments have appropriate types. Consequently, a runtime call to a variadic function that passes inappropriate arguments yields undefined behavior. Such undefined behavior could be exploited to run arbitrary code.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that visits all function
and member function references checking the function declaration for arguments of variadic type. Those defined functions with variadic arguments flag violations of this rule.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}
#include <cstdarg>

char *concatenate(char const *s, ...)
{
  return 0;
}

int main()
{
  char *separator = "\t";
  char *t = concatenate("hello", separator, "world", 0);

  return 0;
}
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution uses a chain of string binary operations instead of a
variadic function.

\begin{verbatim}
#include <string>

string separator = /* some reasonable value */;

string s = "hello" + separator + "world";
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal on all function and member function references.
\item For each function reference check the function declaration and existence of function definition.
\item If function definition does not exist then stop check.
\item Else check function declaration arguments for variadic types.
\item Report any violations.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
\htmladdnormallink{DCL33-C. Do not define variadic functions}{https://www.securecoding.cert.org/confluence/display/cplusplus/DCL33-C.+Do+not+define+variadic+functions}
