% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Right Shift Mask}
\label{RightShiftMask::overview}

% write your introduction
Do not assume that a right shift operation is implemented as either an
arithmetic (signed) shift or a logical (unsigned) shift. If E1 in the
expression E1 >> E2 has a signed type and a negative value, the
resulting value is implementation defined and may be either an
arithmetic shift or a logical shift. Also, be careful to avoid
undefined behavior while performing a bitwise shift.
\subsection{Parameter Requirements}

%Write the Parameter specification here.
No Parameter Required.

\subsection{Implementation}

%Details of the implementation go here.
Upon finding a right shift we trace parent pointers up until we find
a bit and operator.  If we find this bitwise and then we return.  If
we make it to the basic block node of the statement we raise an alert. 


\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection
For implementations in which an arithmetic shift is performed and the
sign bit can be propagated as the number is shifted.
\begin{verbatim}
int stringify;
char buf[sizeof("256")];
sprintf(buf, "%u", stringify >> 24);
% write your non-compliant code example

\end{verbatim}
If stringify has the value 0x80000000, stringify >> 24 evaluates to
0xFFFFFF80 and the subsequent call to sprintf() results in a buffer overflow.
\subsection{Compliant Solution}
For bit extraction, make sure to mask off the bits you are not
interested in.
% write your compliant code subsection

\begin{verbatim}
int stringify;
char buf[sizeof("256")];
sprintf(buf, "%u", ((number >> 24) & 0xff));
% write your compliant code example

\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Find the bitwise and
\item climb parent tree
\item alert if parent null or basicblock found.
\end{enumerate}

\subsection{References}

% Write some references
\htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References}
[Dowd 06] Chapter 6, ``C Language Issues''
[ISO/IEC 9899-1999] Section 6.5.7, ``Bitwise shift operators''
[ISO/IEC 03] Section 6.5.7, ``Bitwise shift operators''