% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{CERT-DCL04-A: Explicit Char Sign}
\label{ExplicitCharSign::overview}
``CERT Secure Coding INT07-A'' states
\begin{quote}
The three types {\tt char}, {\tt signed char}, and {\tt unsigned char} are collectively called the character types. Compilers have the latitude to define {\tt char} to have the same range, representation, and behavior as either {\tt signed char} or {\tt unsigned char}. Irrespective of the choice made, {\tt char} is a separate type from the other two and is {\bf not} compatible with either.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal visiting all SgAssignInitializer nodes. If the SgAssignInitializer node is of type {\tt char} and the operand of the node is a SgCastExp of type {\tt int} or is a SgCharVal whose value is negative then flag an error.

\subsection{Non-Compliant Code Example}
This non-compliant example declares a simple {\tt char} type variable.

\begin{verbatim}
#include <stdio.h>

int main()
{
  int n = 200;
  char c1 = 'i';
  char c2 = n;
  char c3 = 200;
  int i = 1000;

  printf( "%c/c2 = %d\n%c/c3 = %d\n", c1, i/c2, c1, i/c3);

  return 0;
}
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution explicitly declares the {\tt char} variables as {\tt unsigned}.

\begin{verbatim}
#include <stdio.h>

int main()
{
  int n = 200;
  char c1 = 'i';
  unsigned char c2 = n;
  unsigned char c3 = 200;
  int i = 1000;

  printf( "%c/c2 = %d\n%c/c3 = %d\n", c1, i/c2, c1, i/c3);

  return 0;
}
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal visiting all SgAssignInitializer nodes.
\item For each SgAssignInitializer node of type {\tt char}, if the rhs operand is either a cast expression with operand of type int or a negative character value then flag an error.
\item Report any violations. 
\end{enumerate}

\subsection{References}
\htmladdnormallink{Secure Coding : INT07-A. Explicitly specify signed or unsigned for character types}{https://www.securecoding.cert.org/confluence/display/seccode/INT07-A.+Only+use+signed+or+unsigned+char+type+for+numeric+values}
