% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Protect Virtual Methods}
\label{ProtectVirtualMethods::overview}

{\it The Elements of C++ Style} item \#107 states
\begin{quote}
Do not expose virtual methods in the public interface of a class. Use a public methods with a similar name to call the protected virtual method.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal that seeks instances of
SgMemberFunctionDeclaration that have the public access modifier and the
virtual function modifier boolean values set to true. Member functions that
match this pattern are flagged as violations.

\subsection{Non-Compliant Code Example}
This non-compliant code contains a virtual function in the public interface of a class.

\begin{verbatim}
class Class
{
  int n;

  public:
    Class() { n = publicVirtualFunction(); } //constructor
    ~Class() {} //Destructor

    virtual int publicVirtualFunction() { return 1; }
}; //class Class
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution protects the virtual function and adds a public accessor to the virtual function.

\begin{verbatim}
class Class
{
  int n;

  protected:
    virtual int protectedVirtualFunction() { return 1; }
  public:
    Class() { n = publicVirtualFunction(); } //constructor
    ~Class() {} //Destructor
    int publicVirtualFunction(){ return protectedVirtualFunction(); }
}; //class Class
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform a simple AST traversal visiting all function declaration nodes.
\item For each function declaration check the ``public'' and ``virtual'' modifiers. If both ``public'' and ``virtual'' modifiers are set then flag violation.
\item Report any violations. 
\end{enumerate}

\subsection{References}

Bumgardner G., Gray A., and Misfeldt T. {\it The Elements of C++ Style}. Cambridge University Press 2004.
