% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{File Read Only Access}
\label{FileReadOnlyAccess::overview}
Attempting to open files for writing on read-only file systems and files causes errors. This checker checks that all standard C/C++ file I/O is read-only. This checker is normally used only internally for compass verify.

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal visiting all function call expressions calling the function {\tt fopen}. The mode parameter to {\tt fopen} is checked to be either ``r'' or ``rb'' only. Any parameter that violates this restriction is flagged as an error. Additionally, all variable declarations of type {\tt fstream} and {\tt ofstream} are forbidden by this checker.

\subsection{Non-Compliant Code Example}

\begin{verbatim}
#include <fstream>
#include <stdio.h>

int main()
{
  FILE *f1 = fopen( "f1.txt", "w" );
  std::fstream f2( "f2.txt", std::ios::app | std::ios::out );

  FILE *f1r = fopen( "f1.txt", "r" );
  std::ifstream f2r( "f2.txt", std::ios::in );

  return 0;
}
\end{verbatim}

\subsection{Compliant Solution}
Do not write to file.

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal on all function call expressions calling the {\tt fopen} function.
\item Check the mode parameter of {\tt fopen} to be either ``r'' or ``rb''; if not, then flag violation.
\item Perform simple AST traversal on all variable declarations of the type {\tt fstream} or {\tt ofstream}. Forbid the use of these declarations and mark them as violations.
\item Report all violations. 
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
