
\section{File Read Only Access}
\label{FileReadOnlyAccess::overview}
Attempting to open files for writing on read-only file systems and files causes errors. This checker checks that all standard C/C++ file I/O is read-only. This checker is normally used only internally for compass verify.

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal visiting all function call expressions calling the function {\tt fopen}. The mode parameter to {\tt fopen} is checked to be either ``r'' or ``rb'' only. Any parameter that violates this restriction is flagged as an error. Additionally, all variable declarations of type {\tt fstream} and {\tt ofstream} are forbidden by this checker.

\subsection{Non-Compliant Code Example}

 \code{ 

\noindent
\ttfamily
\hlstd{}\hlline{\ \ \ \ 1\ }\hlstd{}\hldir{\#include\ $<$fstream$>$\\
}\hlline{\ \ \ \ 2\ }\hldir{}\hlstd{}\hldir{\#include\ $<$stdio.h$>$\\
}\hlline{\ \ \ \ 3\ }\hldir{}\hlstd{\\
}\hlline{\ \ \ \ 4\ }\hlstd{}\hltyp{int\ }\hlstd{main}\hlsym{()\\
}\hlline{\ \ \ \ 5\ }\hlsym{\{\\
}\hlline{\ \ \ \ 6\ }\hlsym{\hlstd{\ \ }}\hlstd{}\hltyp{FILE\ }\hlstd{*f1\ =\ fopen}\hlsym{(\ }\hlstd{}\hlstr{"f1.txt"}\hlstd{,\ }\hlstr{"w"}\hlstd{\ }\hlsym{)}\hlstd{;\\
}\hlline{\ \ \ \ 7\ }\hlstd{\hlstd{\ \ }std::fstream\ f2}\hlsym{(\ }\hlstd{}\hlstr{"f2.txt"}\hlstd{,\ std::ios::app\ \textbar \ std::ios::out\ }\hlsym{)}\hlstd{;\\
}\hlline{\ \ \ \ 8\ }\hlstd{\\
}\hlline{\ \ \ \ 9\ }\hlstd{\hlstd{\ \ }}\hltyp{FILE\ }\hlstd{*f1r\ =\ fopen}\hlsym{(\ }\hlstd{}\hlstr{"f1.txt"}\hlstd{,\ }\hlstr{"r"}\hlstd{\ }\hlsym{)}\hlstd{;\\
}\hlline{\ \ \ 10\ }\hlstd{\hlstd{\ \ }std::ifstream\ f2r}\hlsym{(\ }\hlstd{}\hlstr{"f2.txt"}\hlstd{,\ std::ios::in\ }\hlsym{)}\hlstd{;\\
}\hlline{\ \ \ 11\ }\hlstd{\\
}\hlline{\ \ \ 12\ }\hlstd{\hlstd{\ \ }}\hlkey{return\ }\hlstd{}\hlnum{0}\hlstd{;\\
}\hlline{\ \ \ 13\ }\hlstd{}\hlsym{\}}\hlstd{}\\
\mbox{}\\
\normalfont













}


\subsection{Compliant Solution}
Do not write to file.

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal on all function call expressions calling the {\tt fopen} function.
\item Check the mode parameter of {\tt fopen} to be either ``r'' or ``rb''; if not, then flag violation.
\item Perform simple AST traversal on all variable declarations of the type {\tt fstream} or {\tt ofstream}. Forbid the use of these declarations and mark them as violations.
\item Report all violations. 
\end{enumerate}

\subsection{References}

