% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Default Case}
\label{DefaultCase::overview}
This test checks to ensure each switch statement has a default
option.  It has been noted that                                                                        
 unexpected cases `falling through'' can be a cause of difficult to
 detect bugs.  A default case will catch those cases. 
% write your introduction

\subsection{Parameter Requirements}

%Write the Parameter specification here.
No parameters required.

\subsection{Implementation}

%Details of the implementation go here.
This implementation checks all statements in the basic                                                                                                               
 block of the switch statement, searching for defaults.  Currently
 this implementation may have false positives (ie. A switch with a
 default will raise an alert) on duff's device.  In general Duff's device does not use
 a default.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}

switch(x)
{
case 1:
...
case 2:
...
}

\end{verbatim}

\subsection{Compliant Solution}

% write your compliant code subsection

\begin{verbatim}
switch(x)
{
case 1:
...
case 2:
...
default:
//handle unexpected cases
}

\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Identifies switch statement
\item Reads all statements in it's basic block searching for default
\item if none found, notify message.
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
\htmladdnormallink{Abbreviated Code Inspection Checklist}{http://www.lore.ua.ac.be/Teaching/SE3BAC} Section 11.2.2, Branching''