% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{No Exit In Mpi Code}
\label{NoExitInMpiCode::overview}

``ALE3D Coding Standards \& Style Guide'' item \#19.1 states that
\begin{quote}
{\tt exit()} must never be called from a parallel code. Calling {\tt exit()} from a parallel code will cause the code to deadlock. Even if you can guarantee that every processor will call {\tt exit()} collectively, this can leave some parallel environments in a hung state because MPI resources are not properly cleaned up. 
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked using a simple AST traversal seeking function reference
expressions. These function reference expressions matching a call to the exit()
function between blocks of MPI code (as delimited between MPI Init and MPI
Finalize) are flagged as checker violations.

\subsection{Non-Compliant Code Example}
This trivial non-compliant code calls {\tt exit()} from an MPI block.

\begin{verbatim}
#include <stdlib.h>
#include "mpi.h"

int main( int argc, char **argv )
{
  MPI_Init( &argc, &argv );
    exit(1);
  MPI_Finalize();

  return 0;
} //main()
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution uses {\tt MPI\_Abort()} instead.

\begin{verbatim}
#include <stdlib.h>
#include "mpi.h"

int main( int argc, char **argv )
{
  MPI_Init( &argc, &argv );
    MPI_Abort( MPI_COMM_WORLD, 1 );
  MPI_Finalize();

  return 0;
} //main()
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform a simple AST traversal of nodes that occur between {\tt MPI\_Init()} and {\tt MPI\_Finalize()} blocks.
\item For each node between MPI blocks, if node is call to {\tt exit()} then flag violation.
\item Report any violations. 
\end{enumerate}

\subsection{References}
Arrighi B., Neely R., Reus J. ``ALE3D Coding Standards \& Style Guide'', 2005.
