% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Friend Declaration Modifier}
\label{FriendDeclarationModifier::overview}

{\it The Elements of C++ Style} item \#96 states that
\begin{quote}
Friend declarations are often indicative of poor design because they bypass access restrictions and hide dependencies between classes and functions.
\end{quote}

\subsection{Parameter Requirements}
This checker takes no parameters and inputs source file.

\subsection{Implementation}
This pattern is checked with a simple AST traversal that seeks declaration statements and determines if any use the ``friend'' modifier keyword. Any declaration statements found with the ``friend'' modifier are flagged as violations.

\subsection{Non-Compliant Code Example}
This non-compliant example uses ``friend'' to access private data.

\begin{verbatim}
class Class
{
  int privateData;
  friend int foo( Class & c );

  public:
    Class(){ privateData=0; }
}; //class Class

int foo( Class & c )
{
  return c.privateData + 1;
} //foo( Class & c )
\end{verbatim}

\subsection{Compliant Solution}
The compliant solution simply uses an accessor function instead.

\begin{verbatim}
class Class
{
  int privateData;

  public:
    Class(){ privateData=0; }
    int getPrivateData(){ return privateData; }
}; //class Class

int foo( Class & c )
{
  return c.getPrivateData() + 1;
} //foo( Class & c )
\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Perform simple AST traversal and visit all declaration statement nodes
\item For each declaration statement check the ``friend'' modifier. If ``friend'' modifier is set then flag violation.
\item Report any violations. 
\end{enumerate}

\subsection{References}

Bumgardner G., Gray A., and Misfeldt T. {\it The Elements of C++ Style}. Cambridge University Press 2004.
