% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Cycle Detection}
\label{CycleDetection::overview}

% write your introduction

This checker is a graph based checker. It determines if cycles are present in the crontrol flow graph.

\subsection{Parameter Requirements}

%Write the Parameter specification here.

Parameters for the run must be provided, e.g. is the analysis interprocedural?
This is work in progress.

\subsection{Implementation}

%Details of the implementation go here.
Within the traversal of the graph, the run function of this checker is called.
This function takes the current and previous node as input. Within the run
function, the algorithm checks the successors of the current node and determines
if a node has been seen before during the graph traversal. If so, a possible
cycle has been found. To verify that a cycle exists, we verify that a path
from the successor to the current node exists.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection

\begin{verbatim}

% write your non-compliant code example

\end{verbatim}

\subsection{Compliant Solution}

% write your compliant code subsection

\begin{verbatim}

% write your compliant code example

\end{verbatim}

\subsection{Mitigation Strategies}
\subsubsection{Static Analysis} 

Compliance with this rule can be checked using structural static analysis checkers using the following algorithm:

\begin{enumerate}
\item Write your checker algorithm
\end{enumerate}

\subsection{References}

% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
