% 9.5.07
% This is a sample documentation for Compass in the tex format.
% We restrict the use of tex to the following subset of commands:
%
% \section, \subsection, \subsubsection, \paragraph
% \begin{enumerate} (no-nesting), \begin{quote}, \item
% {\tt ... }, {\bf ...}, {\it ... }
% \htmladdnormallink{}{}
% \begin{verbatim}...\end{verbatim} is reserved for code segments
% ...''
%

\section{Pointer Comparison}
\label{PointerComparison::overview}

This checker tries to find comparison operations involving
pointers. These comparison results are determined by the memory locations of
the objects pointed to in the program's address space. Using $==$ is usually
safe. But using $<$, $<=$,$>$, and $>=$ is error-prone given the
uncertainty of memory locations of objects across executions and platforms. 

\subsection{Parameter Requirements}

%Write the Parameter specification here.
There is not parameter requirement.

\subsection{Implementation}

%Details of the implementation go here.
The checker finds relational operators such as $<$ , $<=$,$>$, and $>=$
and sends out warnings if at least one of their operands is of pointer types.

\subsection{Non-Compliant Code Example}

% write your non-compliant code subsection
Here is the original code piece which inspired this checker. 
It assumes in the ROSE AST, a parent scope node resides in a memory location of smaller
address compared to the address of its descendent scope node. 
This assumption is wrong given the difference of memory management across
platforms or even executions. A right solution is to walk the AST tree to
find out their relationship.
\begin{verbatim}

SgScopeStatement * scope1, * scope2; 
// ....
if (scope1<scope2)
  //
else
  //

\end{verbatim}

\subsection{References}
No reference.
% Write some references
% ex. \htmladdnormallink{ISO/IEC 9899-1999:TC2}{https://www.securecoding.cert.org/confluence/display/seccode/AA.+C+References} Forward, Section 6.9.1, Function definitions''
